Type Refinements for the Working Class

### 2015-11-24 paper.tex
wordsmithing

Our position is that these views of type theory are not in conflict, but rather
merely describe two distinct layers in a single, harmonious system. From this
perspective, formal type theories can do little more than negotiate matters of
grammar, and therefore may serve as a syntactic (linguistic) framework
for mathematical language, being responsible for the management of variable
binding and substitution. On the other hand, the meaning of mathematical
statements shall be specified \emph{behaviorally} in the semantic type
theory.

The types of the semantic theory can then be said to \emph{refine} the types of
the syntactic theory~\cite{harper-davies:2014, harper-duff:2015, harper:2016},
both by placing restrictions on membership and by coarsening equivalence. Thus
far, most developments of behavioral type theory have been built on a
\emph{unityped} syntactic framework, and so the relation to type refinements
has been difficult to see. In this paper, we contribute a full theory of
behavioral refinements over multi-sorted abstract binding trees, a simple
formal type theory~\cite{harper:2016, sterling-morrison:2015}; this hybrid
system allows the deployment of a Nuprl-style type theory over any signature of
sorts and operators.

>>>>

Our position is that these views of type theory are not in conflict, but rather
merely describe two distinct layers in a single, harmonious system. From this
perspective, formal type theories can do little more than negotiate matters of
grammar, and therefore may serve as a linguistic framework for mathematical
expression, being responsible for the management of variable binding and
substitution. On the other hand, mathematical objects find their \emph{meaning}
in the behavioral type theory, and are subject to an extensional equality.

The types of the semantic theory can then be said to \emph{refine} the types of
the syntactic theory~\cite{harper-davies:2014, harper-duff:2015, harper:2016},
both by placing restrictions on membership and by coarsening equivalence. Thus
far, most developments of behavioral type theory have been built on a
\emph{unityped} syntactic framework, and so the relation to type refinements
has been difficult to see. In this paper, we contribute a full theory of
behavioral refinements over multi-sorted abstract binding trees with symbolic
parameters, a simple but interesting formal type theory~\cite{harper:2016,
sterling-morrison:2015}; this hybrid system allows the deployment of a
Nuprl-style type theory over any signature of sorts and operators.
